Crowdsourcing and phylogenetic modelling reveal parrot tool use is not rare

### 2024-01-16 manuscript.Rmd
prepare for submission

All instances of tool use in our crowdsourcing survey met the established
criteria for tool use in the literature: the manipulation of an unattached 
object as an extension of the animal's body to achieve a goal [@Shumaker2011].
However, a striking feature of the dataset is that the 17 additional species
identified in the video survey were exclusively observed using tools for
self-scratching. While parrots in the wild use tools to achieve a variety of 
goals, including extractive foraging and preening [@Goodman2018;@OHara2021],
parrots in captivity only require tools to achieve the latter goal. Despite this
difference in the type of tool use observed, self-scratching is nonetheless a 
major category of animal tool use. Self-scratching also meets the definition for
a more complex type of embodied tool use known as 'tooling', the deliberate 
generation of a mechanical interface by using an object to manipulate another 
target or surface [@Fragaszy2018]. The fact that we found strong phylogenetic 
clustering of self-scratching tool use from the video survey with other examples
of tool use from the literature supports a common underlying cognitive
mechanism. In line with this, some of the species of parrots that use tools for 
self-scratching in captivity also use tools for other purposes in the wild
[@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021].

>>>>

All instances of tool use in our crowdsourcing survey met the established
criteria for tool use in the literature: the manipulation of an unattached 
object as an extension of the animal's body to achieve a goal [@Shumaker2011].
However, a striking feature of the dataset is that the 17 additional species
identified in the video survey were exclusively observed using tools for
self-scratching. While parrots in the wild use tools to achieve a variety of 
goals, including extractive foraging and preening [@Goodman2018;@OHara2021],
parrots in captivity only require tools to achieve the latter goal. Despite this
difference in the type of tool use observed, self-scratching is nonetheless a 
major category of animal tool use. Self-scratching also meets the definition for
a more complex type of embodied tool use known as 'tooling', the deliberate 
generation of a mechanical interface by using an object to manipulate another 
target or surface [@Fragaszy2018]. This makes it difficult to dismiss
self-scratching as unimportant: to do so would require a redefinition of
tool-use in non-human animals. Moreover, the fact that we found strong 
phylogenetic clustering of self-scratching tool use from the video survey with 
other examples of tool use from the literature supports a common underlying 
cognitive mechanism. In line with this, some of the species of parrots that use
tools for self-scratching in captivity also use tools for other purposes in the 
wild [@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021].

### 2024-01-16 manuscript.tex
prepare for submission

All instances of tool use in our crowdsourcing survey met the established
criteria for tool use in the literature: the manipulation of an unattached
object as an extension of the animal's body to achieve a goal\textsuperscript{10}.
However, a striking feature of the dataset is that the 17 additional species
identified in the video survey were exclusively observed using tools for
self-scratching. While parrots in the wild use tools to achieve a variety of
goals, including extractive foraging and preening\textsuperscript{34,38},
parrots in captivity only require tools to achieve the latter goal. Despite this
difference in the type of tool use observed, self-scratching is nonetheless a
major category of animal tool use. Self-scratching also meets the definition for
a more complex type of embodied tool use known as `tooling', the deliberate
generation of a mechanical interface by using an object to manipulate another
target or surface\textsuperscript{65}. The fact that we found strong phylogenetic
clustering of self-scratching tool use from the video survey with other examples
of tool use from the literature supports a common underlying cognitive
mechanism. In line with this, some of the species of parrots that use tools for
self-scratching in captivity also use tools for other purposes in the wild\textsuperscript{30,31,34,38}.

>>>>

All instances of tool use in our crowdsourcing survey met the established
criteria for tool use in the literature: the manipulation of an unattached
object as an extension of the animal's body to achieve a goal\textsuperscript{10}.
However, a striking feature of the dataset is that the 17 additional species
identified in the video survey were exclusively observed using tools for
self-scratching. While parrots in the wild use tools to achieve a variety of
goals, including extractive foraging and preening\textsuperscript{34,38},
parrots in captivity only require tools to achieve the latter goal. Despite this
difference in the type of tool use observed, self-scratching is nonetheless a
major category of animal tool use. Self-scratching also meets the definition for
a more complex type of embodied tool use known as `tooling', the deliberate
generation of a mechanical interface by using an object to manipulate another
target or surface\textsuperscript{65}. This makes it difficult to dismiss
self-scratching as unimportant: to do so would require a redefinition of
tool-use in non-human animals. Moreover, the fact that we found strong
phylogenetic clustering of self-scratching tool use from the video survey with
other examples of tool use from the literature supports a common underlying
cognitive mechanism. In line with this, some of the species of parrots that use
tools for self-scratching in captivity also use tools for other purposes in the
wild\textsuperscript{30,31,34,38}.
